---
title: "Coursera - Regression Models"
author: "Enlik"
date: "19 February 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary
This analysis using dataset comes from `mtcars` package. The data was extracted from the 1974 Motor Trend US magazine and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). The main purpose of this analysis is to answer the following two questions:

1. Is an automatic or manual transmission better for MPG
2. Quantify the MPG difference between automatic and manual transmissions


Based on the data, this document shows that the manual transmission is better than the automatic for the mpg: all the models tested say this.

The best model says that the manual transmission allows doing an average of 2.396 miles plus then the automatic.
    

# Read Data
```{r}
data(mtcars)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)

str(mtcars)
summary(mtcars)
```
mtcars dataset have 32 observations and 11 variables:

1. mpg: Miles/(US) gallon
2. cyl: Number of cylinders
3. disp: Displacement (cu.in.)
4. hp: Gross horsepower
5. drat: Rear axle ratio
6. wt: Weight (1000 lbs)
7. qsec: 1/4 mile time
8. vs: V/S
9. am: Transmission (0 = automatic, 1 = manual)
10. gear: Number of forward gears
11. carb: Number of carburetors


```{r}
boxplot(mpg ~ am, data = mtcars,
         col  = c("green", "yellow"),
         xlab = "Transmission",
         ylab = "Miles per Gallon (MPG)",
         main = "MPG by Transmission Type",
         names= c("automatic","manual"),
         horizontal= T) 
```

From boxplot graph above, it seems manual transmission was better than automatic.


```{r}
auto=subset(mtcars,select=mpg,am==0)
manual=subset(mtcars,select=mpg,am==1)
t.test(auto,manual)
```

# Simple Linear Regression
```{r}
regSIM <- lm(mpg~am,mtcars) 
summary(regSIM) 
```
R-square = 0.36 means the model have 36% of the variance.

# Multivariate Regression
```{r}
regTOT <- lm(mpg~.,mtcars)
summary(regTOT)
```

I use the stepwise regression method for choice the best variables for explain the mpg values.
```{r}
regSR=step(regTOT,trace=0)
summary(regSR)
```
It explains more about mpg values that are wt, qsec and am.
This model has 84.97% of the variance and has all the coefficients significative at 5%, so this model is better than the other two.
For this model the manual transmission allow to do 2.396 miles plus then the automatic.

```{r}
anova(regSIM,regSR,regTOT)
```

Anova confirm that the model with 3 predictor (wt, qsec, am), is the best choice.

# Residuals
Residual plot of the best model:
```{r}
plot(regSR, which=c(1:1))
```

# Conclusion
Based on the data, this document shows that the manual transmission is better than the automatic for the mpg: all the models tested say this.

The best model says that the manual transmission allows doing an average of 2.396 miles plus then the automatic.
